---
title: Car Colors Over Time
author: ''
date: '2020-10-11'
slug: car-colors-over-time
categories: []
tags: []
---

# HMB (Hold my baby...)

This is my first attempt at quick plot beautification. "Hold my beer" is a punchline (or lead up) to many a joke in the US with often a picture or graphic at the end of the joke to demonstrate without words what happened. I don't like beer, but I do have a baby that demands me very frequently. This is an attempt of my self care for my brain during these early months.

## Car Colors

My husband showed me a [graphic of car colors](https://www.reddit.com/r/dataisbeautiful/comments/j7yfqb/oc_car_color_distribution_broken_down_by/) over time that I thought was fun, but I also thought I could make it prettier. 

A quick Google source lead me here to see the post he had shown me and here is where I grabbed the [OP's data](https://pastebin.com/1gb8vAAr). 

```{r libraries, include=FALSE}
library(tidyverse)
library(lubridate)
library(plyr)
library(ggthemes)
```
```{r import_clean_data}

car_colors <- read.delim2("~/Tidy_TidyTuesday/data/car_color.txt")

car_long <- pivot_longer(car_colors, 2:14, names_to = "color", values_to = "percentage") 

car_long$percentage <- sub("%", "", car_long$percentage)

car_long$percentage <- sub(",", ".", car_long$percentage)

car_long$percentage <- as.numeric(car_long$percentage)

car_long$color <- as_factor(car_long$color) 

color_names <- levels(car_long$color)

car_long$color <- revalue(car_long$color, c("Biege" = "Beige"))

car_long$color <- factor(car_long$color, levels = c(
  "Purple",
    "Blue",
    "Green",
    "Yellow",
    "Gold",
    "Beige",
    "Brown" , 
    "Red" , 
    "Burgundy" , 
    "White" ,
    "Silver" ,
    "Gray" ,
    "Black" )
)
```

## Plot this thang!

```{r}
plot <- ggplot(car_long, aes(x = Year, y = percentage)) +
  geom_col(aes(fill = color)) +
  scale_fill_manual(values = c(
    "Purple" = "#800080",
    "Blue" = "#0000ff",
    "Green" = "#008000",
    "Yellow" = "#ffff00",
    "Gold" = "#ffd700",
    "Beige" = "#f5f5dc",
    "Brown" = "#a52a2a", 
    "Red" = "#ff0000", 
    "Burgundy" = "#800020", 
    "White" = "#ffffff",
    "Silver" = "#c0c0c0",
    "Gray" = "#808080",
    "Black" = "#000000")
  ) 

plot +
    labs(title = "Colors of Cars Over Time in Poland", subtitle = "Mara's quick redo", x = "Production Year", y = "Percentage", caption = "OP: https://tinyurl.com/y4hhlmbc") + 
  theme_economist() +
  theme(legend.position = "none")
```
